---
title: "`easyrbi`: A Package for retrieving USGS gauges station and landuse data, and calculating stream flashiness"
author: "Wayne Ndlovu and Annah Mutaya"
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{`easyrbi`: A Package for retrieving USGS gauges station and landuse data, and calculating stream flashiness}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
tags:
  - R
  - Rstats
  - tidyverse
  - dataRetrieval
  - easyrbi
authors:
  - name: Wayne Ndlovu
  - name: Annah Mutaya
output:
  rmarkdown::html_vignette:
    df_print: default
    number_sections: no
---
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(dplyr)
library(tidyr)
library(purrr)
library(dataRetrieval)
library(easyrbi)
library(trend)
library(Kendall)
library(usethis)
```


## Summary
The `easyrbi` package is made with users and hydrologists who are interested in studying stream discharge data in mind. The package contains functions that retrieve streamflow data (pCode = 00060) from the [USGS website](https://waterservices.usgs.gov/rest/Site-Test-Tool.html) and dam removal time series data from [USGS Gages II](https://www.sciencebase.gov/catalog/item/59692a64e4b0d1f9f05fbd39) collected between 1931-01-01 and 2014-12-31.


## Introduction
To get started, install the `easyrbi` package
```{r, eval=FALSE, message=FALSE}
# If you haven't installed remotes yet, do so:
# install.packages("remotes")
remotes::install_github("amutaya/easyrbi")
```



anna - rbi_df, site_data, trends
wayne - basin id, dam_removal, dam_trends


#### Investigate the trends at sites where dams were removed.
- For any given gauge stations, check if there is a relationship between trends and dam removals 
```{r, eval=TRUE, warning=FALSE, message=FALSE}
dam_trends(c("01092000", "01100000", "01208500"), "1970-10-01", "1990-09-30")
```
