---
title: "`easyrbi`: A Package for retrieving USGS gauges station and landuse data, and calculating stream flashiness"
author: "Wayne Ndlovu and Annah Mutaya"
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{`easyrbi`: A Package for retrieving USGS gauges station and landuse data, and calculating stream flashiness}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
tags:
  - R
  - Rstats
  - tidyverse
  - dataRetrieval
  - easyrbi
authors:
  - name: Wayne Ndlovu
  - name: Annah Mutaya
output:
  rmarkdown::html_vignette:
    df_print: default
    number_sections: no
---
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(dplyr)
library(tidyr)
library(purrr)
library(dataRetrieval)
library(easyrbi)
library(trend)
library(Kendall)
library(usethis)
```


## Summary
The `easyrbi` package is made with users and hydrologists who are interested in studying stream discharge data in mind. The package contains functions that retrieve streamflow data (pCode = 00060) from the [USGS website](https://waterservices.usgs.gov/rest/Site-Test-Tool.html) and dam removal time series data from [USGS Gages II](https://www.sciencebase.gov/catalog/item/59692a64e4b0d1f9f05fbd39) collected between 1931-01-01 and 2014-12-31. All functions return tidy dataframes which are easy to work with.


## Introduction
To get started, install the `easyrbi` package and load all the necessary packages
```{r, eval=FALSE, message=FALSE}
# If you haven't installed remotes yet, do so:
# install.packages("remotes")
remotes::install_github("amutaya/easyrbi")

library(easyrbi)
library(tidyverse)
library(trend)
library(Kendall)
library(dataRetrieval)
library(usethis)

# create sample sites
study_sites <- c("01564500", "01567000", "01567500", "01568000", "01570000", "01571500", "01573000", "01574000", "01574500", "01576500")
```


## Uses

### Watershed Identification
Ever wondered how large the watershed you are studying is? Or whether the watershed is consired a reference or non-refence site? For any given site numbers, the `basin_id` function generates a dataframe that consists of the watershed identification information as provided by the U.S. Geological Survey (USGS) Geospatial Attributes of Gages for Evaluating Streamflow II (GAGES-II) dataset. The `basin_id` function is interactive, so you can choose to overwrite previously downloaded data files.

#### Example 1
```{r, eval=TRUE, warning=FALSE, message=FALSE}
basin_id(c("01567000", "01490000", "01492500"))
```

#### Example 2
```{r, eval=TRUE, warning=FALSE, message=FALSE}
#basin_id(study_sites)
```

### How many dams have been removed close to your study sites?
`easyrbi` allows the user to also download dam removal data recorded between 1931 and 2014 using a simple function that only takes in the site number(s) as an argument. The `dam_removal` function is also interactive, so you can choose to overwrite previously downloaded data files.

#### Example 1
```{r, eval=TRUE, warning=FALSE, message=FALSE}
#dam_removal("01066000")
```

#### Example 2
```{r, eval=TRUE, warning=FALSE, message=FALSE}
#dam_removal(study_sites)
```

### Investigate the trends at sites where dams were removed.
The `easyrbi` package also help you get a better understanding of the impacts of dam removals on stream flashiness. For any given gauge stations, check if there is a relationship between stream flashiness trends and dam removals.

#### Example 1
```{r, eval=TRUE, warning=FALSE, message=FALSE}
#dam_trends(c("01092000", "01100000", "01208500"), "1970-10-01", "1990-09-30")
```

#### Example 2
```{r}
#dam_trends(study_sites)
```

